The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-03 manuscript.Rmd
revise relation of Rocker and repo2docker in WT
Co-authored-by: Dirk Eddelbuettel <edd@debian.org>

Second, the **Whole&nbsp;Tale** project (https://wholetale.org) combines the strengths of the Rocker Project's curated Docker images with `repo2docker`.
Whole&nbsp;Tale is a National Science Foundation (NSF) funded project developing a scalable, open source, multi-user platform for reproducible research \citep{brinckman2019, chard2019a}.
A central goal of the platform is to enable researchers to easily create and publish executable research objects\footnote{In Whole~Tale a \emph{tale} is a research object that contains metadata, data (by copy or reference), code, narrative, documentation, provenance, and information about the computational environment to support computational reproducibility.} associated with published research \citep{chard2019b}.
Using Whole&nbsp;Tale, researchers can create and publish Rocker-based reproducible research objects to a growing number of repositories including DataONE member nodes, Zenodo and soon Dataverse.
Additionally, Whole&nbsp;Tale supports automatic data citation and is working on capabilities for image preservation and provenance capture to improve the transparency of published computational research artefacts \citep{mecum2018, mcphillips2019}.
For R users, Whole&nbsp;Tale extends the Jupyter Project's `repo2docker` component to simplify the customisation of R-based environments for researchers with limited experience with either Docker or Git.
Multiple options have been discussed to allow users to change the base image used in `repo2docker` from the default Ubuntu LTS (long-term support) required to support the Rocker Project images.
Whole&nbsp;Tale implemented a custom `RockerBuildPack`\footnote{See \href{https://github.com/whole-tale/repo2docker\_wholetale}{https://github.com/whole-tale/repo2docker\_wholetale}} to support customisation of the `rocker/geospatial` image through `repo2docker` composability\footnote{Composability refers to the ability to combine multiple package managers -- such as R, `pip`, and `conda`}.
This works in part because Rocker images are based on a Debian distribution, so the instructions created by `repo2docker` for Ubuntu work because of compatible shell and package manager.

>>>>

Second, the **Whole&nbsp;Tale** project (https://wholetale.org) combines the strengths of the Rocker Project's curated Docker images with `repo2docker`.
Whole&nbsp;Tale is a National Science Foundation (NSF) funded project developing a scalable, open source, multi-user platform for reproducible research \citep{brinckman2019, chard2019a}.
A central goal of the platform is to enable researchers to easily create and publish executable research objects\footnote{In Whole~Tale a \emph{tale} is a research object that contains metadata, data (by copy or reference), code, narrative, documentation, provenance, and information about the computational environment to support computational reproducibility.} associated with published research \citep{chard2019b}.
Using Whole&nbsp;Tale, researchers can create and publish Rocker-based reproducible research objects to a growing number of repositories including DataONE member nodes, Zenodo and soon Dataverse.
Additionally, Whole&nbsp;Tale supports automatic data citation and is working on capabilities for image preservation and provenance capture to improve the transparency of published computational research artefacts \citep{mecum2018, mcphillips2019}.
For R users, Whole&nbsp;Tale extends the Jupyter Project's `repo2docker` tool to simplify the customisation of R-based environments for researchers with limited experience with either Docker or Git.
Multiple options have been discussed to allow users to change the Ubuntu LTS (long-term support, currently _Bionic Beaver_) base image, `buildpack-deps:bionic`, used in `repo2docker`.
Whole&nbsp;Tale implemented a custom `RockerBuildPack`\footnote{See \href{https://github.com/whole-tale/repo2docker\_wholetale}{https://github.com/whole-tale/repo2docker\_wholetale}.}.
The build pack combines a `rocker/geospatial` image with `repo2docker`'s composability\footnote{Composability refers to the ability to combine multiple package managers and their configuration files, such as R, `pip`, and `conda`; see Section~\nameref{binder} for details.}.
This works because both Rocker images and the `repo2docker` base image use distributions with APT \citep{wikipedia_contributors_apt_2020} so that the instructions created by the latter work because of the compatible shell and package manager.
